---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---


## Loading and preprocessing the data

Reads the data and saves it to a table called activity.

```{r echo=TRUE}
setwd("C:/RepData_PeerAssessment1") 
unzip("activity.zip")  
activity<-read.csv("activity.csv", header=TRUE) 
```



## What is mean total number of steps taken per day?

First it's calculated the sum of steps by day and after it's ploted in a histogram

```{r}
stepsday<-rowsum(activity$steps,activity$date,na.rm = T)
hist(stepsday, xlab = "Steps by day",main = "Histogram of steps by day")

meanst<-round(mean(stepsday))
medianst<-round(median(stepsday))
```

The mean of steps taken per day is `r meanst` and the median is `r as.character(medianst)`


## What is the average daily activity pattern?

This is a 5-minute interval time series plot vs the mean of steps for each interval

```{r}
stepsinterval<-tapply(activity$steps,activity$interval,mean, na.rm=TRUE)
plot(unique(activity$interval),stepsinterval,type = "l", xlab = "Intervals",
     ylab = "Steps")
```

This is the interval with the maximum number of steps

```{r}
stepsinterval[which.max(stepsinterval)]
```


## Imputing missing values

First it's calculated the number of NAs in the dataset

```{r}
sum(is.na(activity$steps))
```

Then each NA is replaced by the mean in the interval. 

So for example, if interval 0 on 2012-10-01 is NA then it is replaced by the mean of all the intervals 0, previously calculated

```{r}
stepsint<-stepsinterval
intervs<-unique(activity$interval)

setinterval<-data.frame(intervs,stepsint)
activity2<-activity

for (i in 1:length(activity2$steps))
{
    if(is.na(activity2$steps[i]))
    {interv<-activity2$interval[i]
    activity2$steps[i]=setinterval$stepsint[which(setinterval$intervs==interv)]
    activity2$steps=round(activity2$steps)
    }
}
```

Hence this histogram is generated

```{r}
stepsday2<-rowsum(activity2$steps,activity2$date)
hist(stepsday2, xlab = "Steps by day",main = "Histogram of steps by day")

meanst2<-round(mean(stepsday))
medianst2<-round(median(stepsday))
```

The mean of steps taken per day with NAs replaced is `r meanst2` and the median is `r as.character(medianst2)`

## Are there differences in activity patterns between weekdays and weekends?

For clasification purposes, a weekdays column is created. Setlocale is used to set the days of the week in English.

```{r}
Sys.setlocale("LC_ALL","English")
activity2$date<-as.POSIXct(activity2$date)
activity2$weekdays<-weekdays(activity2$date)
```

Then, using this column, another column is created to clasify as Weekday or Weekend

```{r}
days<-c("Monday","Tuesday","Wednesday","Thursday","Friday")
for (i in 1:length(activity2$weekdays))
{
  if(activity2$weekdays[i]%in%days)
  {
      activity2$state[i]<-"Weekday"
  }
    else
    {
        activity2$state[i]<-"Weekend"
    }
}
```

Using this, two new datasets are created. One for weekdays, one for weekend.

```{r}
weekdaysst<-activity2[which(activity2$state=="Weekday"),]
weekendst<-activity2[which(activity2$state=="Weekend"),]
```

And both of them are plotted.

```{r}
library(lattice)

stweekday<-tapply(weekdaysst$steps,weekdaysst$interval,mean)
xyplot(stweekday ~ unique(interval), data = activity2, type = "l",
       xlab = "Interval", ylab = "Steps")

stweekend<-tapply(weekendst$steps,weekendst$interval,mean)
xyplot(stweekend ~ unique(interval), data = activity2, type = "l", 
      xlab = "Interval", ylab = "Steps")
```

As we can see there are differences in the weekends.Especially in the interval of 10:00 and 20:00.